An open library of human kinase domain constructs for automated bacterial expression

### 2016-02-02 manuscript/ms.tex
clarified/fix a few things

The protein databank (PDB) now contains over 100 human kinases that---according to the PDB data records---were expressed in bacteria.
Since bacterial expression is often complicated by the need to tailor expression and purification protocols individually for each protein expressed, we wondered whether a simple, uniform, automatable expression and purification protocol could be used to express a large number of human kinases to produce a convenient bacterial expression library to facilitate kinase research and selective inhibitor development.
As a first step toward this goal, we developed a structural informatics pipeline to use available kinase structural data and associated metadata to select constructs from available human kinase libraries to clone into a standard set of vectors intended for phosphatase coexpression.
Automated expression screening in Rosetta2 cells found that 52 human kinase domains express with yields greater than 2 mg/L, which should be usable for biochemical, biophysical, screening, and structural biology studies.

@@

A panel containing the 96 kinase domain constructs selected through our semi-automated method, was tested for expression in \emph{E. coli}. 
From this initial test, 52 kinase domains showed reasonable expression (yield of more than 2 mg/L in the eluate) (Table~\ref{expression_table}).
While the initial panel of 96 kinases was well-distributed across kinase families, the final most highly expressing (yield of more than 100 ng/$\mu$l eluate) were not as evenly distributed (Figure~\ref{fig:kinome_expression_tree}). 
The 17 most highly expressing kinases showed relatively high purity after elution, though we note that eluting via TEV site cleavage results in a quantity of TEV protease in the eluate (Figure~\ref{fig:caliper_image}).

@@

\caption{{\bf Expression results by kinase.} Yield (determined by Caliper GX II quantitation of the expected size band) reported in ng/$\mu$l eluate, where total eluate volume was 120 $\mu$l from 900 $\mu$L bacterial culture.}

@@

\begin{figure*}[tb]
   \includegraphics[width=\columnwidth]{figures/caliper_image.png}
  \caption{{\bf Synthetic gel image rendering of highest expressing kinases.}
   Caliper GX II synthetic gel image rendering of kinases expressing > 25 mg/L from microfluidic capillary electrophoresis quantitation.
   }
  \label{fig:caliper_image}
\end{figure*}

>>>>

The protein databank (PDB) now contains over 100 human kinases that---according to the PDB data records---were expressed in bacteria.
Since bacterial expression is often complicated by the need to tailor expression and purification protocols individually for each protein expressed, we wondered whether a simple, uniform, automatable expression and purification protocol could be used to express a large number of human kinases to produce a convenient bacterial expression library to facilitate kinase research and selective inhibitor development.
As a first step toward this goal, we developed a structural informatics pipeline to use available kinase structural data and associated metadata to select constructs from available human kinase libraries to clone into a standard set of vectors intended for phosphatase coexpression.
Automated expression screening in Rosetta2 cells found that 52 human kinase domains express with yields greater than 2 mg/L culture, which should be usable for biochemical, biophysical, screening, and structural biology studies.

@@

A panel containing the 96 kinase domain constructs selected through our semi-automated method, was tested for expression in \emph{E. coli}. 
From this initial test, 52 kinase domains showed reasonable expression (yield of more than 2 ng/$\mu$L eluate, which corresponds to 2 mg/L culture) (Table~\ref{expression_table}).
While the initial panel of 96 kinases was well-distributed across kinase families, the final most highly expressing (yield of more than 12 mg/L kinase) were not as evenly distributed (Figure~\ref{fig:kinome_expression_tree}). 
The 17 most highly expressing kinases showed relatively high purity after elution, though we note that eluting via TEV site cleavage results in a quantity of TEV protease in the eluate (Figure~\ref{fig:caliper_image}).

@@

\caption{{\bf Expression results by kinase.} Yield (determined by Caliper GX II quantitation of the expected size band) reported in mg/L culture, where total eluate volume was 120 $\mu$l from 900 $\mu$L bacterial culture.}

@@

\begin{figure*}[tb]
   \includegraphics[width=\columnwidth]{figures/caliper_image.png}
  \caption{{\bf Synthetic gel image rendering of highest expressing kinases.}
   Caliper GX II synthetic gel image rendering of kinases expressing > 25 mg/L culture from microfluidic capillary electrophoresis quantitation.
   }
  \label{fig:caliper_image}
\end{figure*}
